Anony\emph{mice}d shareable data: Using \emph{mice} to create and analyze multiply imputed synthetic data sets

### 2021-11-01 manuscript/Manuscript.Rmd
update Gerko

---
title: "Anony*mice*d shareable data: Using *mice* to create and analyze multiply imputed synthetic data sets"
author:
  - name: Thom Volker
    affil: 1,*, \dagger
    orcid: 0000-0002-2408-7820
  - name: Gerko Vink
    affil: 1, \dagger
    orcid: 0000-0001-9767-1924
affiliation:
  - num: 1
    address: |
      Utrecht University - 
      Department of Methodology and Statistics. 
      Padualaan 14. 3584CH Utrecht, the Netherlands
# firstnote to eighthnote
firstnote: |
  These authors contributed equally to this work.
correspondence: |
  t.b.volker@uu.nl.
journal: psych
type: article
status: submit
bibliography: synth.bib
appendix: appendix.tex
abstract: |
  Synthetic data sets can greatly improve the dissemination of data 
  and further analysis of private data. Generating and analyzing synthetic data 
  sets is straightforward, yet a synthetic data analysis pipeline is 
  seldom adopted by applied researchers. We outline a simple procedure for 
  generating and analyzing synthetic data sets with `mice` in `R`. We demonstrate 
  using simulations that the analysis results obtained on synthetic data 
  yields unbiased and valid inferences, and leads to synthetic 
  records that cannot be distinguished from the true data records. The ease of use 
  when synthesizing data with `mice`, together with the validity of inferences 
  obtained through this procedure opens up a wealth of possibilities for data 
  dissemination and further research of initially private data. 
keywords: |
  mice; imputation; synthetic data.
acknowledgement: |
  We are grateful to Mirthe Hendriks and Stijn van den Broek for replicating the 
  simulation study on an independent data set. Additionally, we thank Hanne Oberman and two anonymous reviewers for their stimulating comments.
authorcontributions: |
  The authors contributed equally to this work.
conflictsofinterest: |
  The authors declare no conflict of interest. 
sampleavailability: |
  A full simulation archive is available from https://github.com/amices/Federated_imputation/tree/master/manuscript.
output: rticles::mdpi_article
---

@@

Anonymizing research data might seem a quick and appealing approach to limit the unique identification of participants. However, this approach is not sufficient to fulfil contemporary privacy and confidentiality requirements [@ohm_broken_2009; @national_putting_2007]. Over the years, several other techniques have been used to increase the confidentiality of research data, such as categorizing continuous variables, top coding values above an upper bound or adding random noise to the observed values [@drechsler_synthetic_2011]. However, these methods may distort the true data relation between variables, thereby reducing the data quality and the scientific returns for re-using the same data for further research.

@@

Generally, there are three ways to impute data: joint modeling, sequential modeling and fully conditional specification [@murray_multiple_2018; @ludtke2020regression; @grund2021using]. With joint modeling, the entire joint distribution of the data is specified as a single multivariate distribution, and the imputations are drawn from this model. This strategy is implemented in, among others, the `R`-package `jomo` [@jomo] to deal with missing data. However, when the structure of the data increases in complexity, specifying a single multivariate distribution can become cumbersome. A more flexible solution has been proposed in sequential regression imputation, in which the multivariate distribution is separated into univariate distributions for every variable. Every single variable is then imputed conditional on a model containing only the variables in the sequence located before the variable to be imputed. This approach has been implemented in the R-packages `mdmb` [@mdmb] for imputation of missing data and `synthpop` [@synthpop] for imputation of synthetic data.

@@

```{r eval = FALSE}
library(mice)
imp <- mice(data.to.synthesize, 
            m = 10,
            maxit = 1,
            where = matrix(TRUE,
                           nrow = nrow(data.to.synthesize),
                           ncol = ncol(data.to.synthetize)))
```

@@

However, one of the methods that has shown a reasonable performance over multiple simulations uses FCS in combination with classification and regression trees [CART\; @breiman_cart_1984; @reiter_cart_2005; @burgette_reiter_cart_2010; @doove_buuren_recursive_2014; @raab_practical_2016]. Due to its non-parametric nature, CART allows to model more complex patterns (e.g., interactions and non-linear relationships) in the data. Loosely speaking, CART sequentially splits the predictor space into non-overlapping regions in such a way that the within-region variance is as small as possible after every split. As such, CART does not impose any parametric distribution on the data, making it a widely applicable method that allows for a large variety of relationships within the data [@islr_2013]. Given these appealing characteristics and the call for the use of flexible methods when multiply imputing data, we will focus our illustrations and evaluations of `mice` to method `mice.impute.cart()`, realized by:

@@

In a nutshell, the above code shows the simplicity of creating $m = 10$ synthetic data sets using `mice`. In practice, however, one should take some additional complicating factors into account. For example, one should account for deterministic relations in the data. Additionally, relations between variables may be described best using a different model than `CART`. Such factors are data dependent and should be considered by the imputer. In the next section, we will describe how the `mice::boys` data [fredriks_boys_2000] can be adequately synthesized. Additionally, we will show through simulations that this approach yields valid inferences.

@@

We see from Table 1 that the synthetic data estimates closely resemble the true data estimates. All sample statistics of interest show negligible bias over the 1000 synthetic data sets. Hence, univariately the imputation model proves adequate.

>>>>

---
title: "Anony*mice*d shareable data: Using *mice* to create and analyze multiply imputed synthetic data sets"
author:
  - name: Thom Volker
    affil: 1,*, \dagger
    orcid: 0000-0002-2408-7820
  - name: Gerko Vink
    affil: 1, \dagger
    orcid: 0000-0001-9767-1924
affiliation:
  - num: 1
    address: |
      Utrecht University - 
      Department of Methodology and Statistics. 
      Padualaan 14. 3584CH Utrecht, the Netherlands
# firstnote to eighthnote
firstnote: |
  These authors contributed equally to this work.
correspondence: |
  t.b.volker@uu.nl.
journal: psych
type: article
status: submit
bibliography: synth.bib
appendix: appendix.tex
abstract: |
  Synthetic data sets can greatly improve the dissemination of data 
  and further analysis of private data. Generating and analyzing synthetic data 
  sets is straightforward, yet a synthetic data analysis pipeline is 
  seldom adopted by applied researchers. We outline a simple procedure for 
  generating and analyzing synthetic data sets with `mice` in `R`. We demonstrate 
  using simulations that the analysis results obtained on synthetic data 
  yields unbiased and valid inferences, and leads to synthetic 
  records that cannot be distinguished from the true data records. The ease of use 
  when synthesizing data with `mice`, together with the validity of inferences 
  obtained through this procedure opens up a wealth of possibilities for data 
  dissemination and further research of initially private data. 
keywords: |
  mice; imputation; synthetic data.
acknowledgement: |
  We are grateful to Mirthe Hendriks and Stijn van den Broek for replicating the 
  simulation study on an independent data set. Additionally, we thank Hanne Oberman and two anonymous reviewers for their helpful and insightful comments.
authorcontributions: |
  The authors contributed equally to this work.
conflictsofinterest: |
  The authors declare no conflict of interest. 
sampleavailability: |
  A full simulation archive is available from https://github.com/amices/Federated_imputation/tree/master/manuscript.
output: rticles::mdpi_article
---

@@

Anonymizing research data might seem a quick and appealing approach to limit the unique identification of participants. However, this approach is not sufficient to fulfill contemporary privacy and confidentiality requirements [@ohm_broken_2009; @national_putting_2007]. Over the years, several other techniques have been used to increase the confidentiality of research data, such as categorizing continuous variables, top coding values above an upper bound or adding random noise to the observed values [@drechsler_synthetic_2011]. However, these methods may distort the true data relation between variables, thereby reducing the data quality and the scientific returns for re-using the same data for further research.

@@

Generally, there are three ways to impute data: joint modeling, sequential modeling and fully conditional specification [@murray_multiple_2018; @ludtke2020regression; @grund2021using]. With joint modeling, the entire joint distribution of the data is specified as a single multivariate distribution, and the imputations are drawn from this model. This strategy is implemented in, among others, the `R`-package `jomo` [@jomo] to deal with missing data. However, when the structure of the data increases in complexity, specifying a single multivariate distribution that would fit the observed data can become challenging. A more flexible solution has been proposed in sequential regression imputation, in which the multivariate distribution is separated into univariate conditional distributions. Every single variable is then imputed conditionally on a model containing only the variables in the sequence located before the variable to be imputed. This approach has been implemented in the R-packages `mdmb` [@mdmb] for imputation of missing data and `synthpop` [@synthpop] for imputation of synthetic data.

@@

```{r eval = FALSE}
library(mice)
syn <- mice(data.to.synthesize, 
            m = 10,
            maxit = 1,
            where = matrix(TRUE,
                           nrow = nrow(data.to.synthesize),
                           ncol = ncol(data.to.synthetize)))
```

@@

However, one of the methods that has shown a reasonable and consistent performance over multiple simulations using FCS in combination with classification and regression trees [CART\; @breiman_cart_1984; @reiter_cart_2005; @burgette_reiter_cart_2010; @doove_buuren_recursive_2014; @raab_practical_2016]. Due to its non-parametric nature, CART allows to model more complex patterns (e.g., interactions and non-linear relationships) in the data. Loosely speaking, CART sequentially splits the predictor space into non-overlapping regions in such a way that the within-region variance is as small as possible after every split. As such, CART does not impose any parametric distribution on the data, making it a widely applicable method that allows for a large variety of relationships within the data [@islr_2013]. Given these appealing characteristics and the call for the use of flexible methods when multiply imputing data, we will focus our illustrations and evaluations of `mice` to method `mice.impute.cart()`, realized by:

@@

In a nutshell, the above code shows the simplicity of creating $m = 10$ synthetic data sets using `mice`. In practice, however, one should take some additional complicating factors into account. For example, one should account for deterministic relations in the data. Additionally, relations between variables may be described best using a different model than `CART`. Such factors are data dependent and should be considered by the imputer. In the next section, we will describe how the `mice::boys` data [@fredriks_boys_2000] can be adequately synthesized. Additionally, we will show through simulations that this approach yields valid inferences.

@@

We see from Table 1 that the synthetic data estimates closely resemble the true data estimates. All sample statistics of interest show negligible bias over the 1000 synthetic data sets. Hence, univariately the synthetisation model proves adequate.

### 2021-11-01 manuscript/Manuscript.tex
update Gerko

Anonymizing research data might seem a quick and appealing approach to
limit the unique identification of participants. However, this approach
is not sufficient to fulfil contemporary privacy and confidentiality
requirements \citep{ohm_broken_2009, national_putting_2007}. Over the
years, several other techniques have been used to increase the
confidentiality of research data, such as categorizing continuous
variables, top coding values above an upper bound or adding random noise
to the observed values \citep{drechsler_synthetic_2011}. However, these
methods may distort the true data relation between variables, thereby
reducing the data quality and the scientific returns for re-using the
same data for further research.

@@

Generally, there are three ways to impute data: joint modeling,
sequential modeling and fully conditional specification
\citep{murray_multiple_2018, ludtke2020regression, grund2021using}. With
joint modeling, the entire joint distribution of the data is specified
as a single multivariate distribution, and the imputations are drawn
from this model. This strategy is implemented in, among others, the
\texttt{R}-package \texttt{jomo} \citep{jomo} to deal with missing data.
However, when the structure of the data increases in complexity,
specifying a single multivariate distribution can become cumbersome. A
more flexible solution has been proposed in sequential regression
imputation, in which the multivariate distribution is separated into
univariate distributions for every variable. Every single variable is
then imputed conditional on a model containing only the variables in the
sequence located before the variable to be imputed. This approach has
been implemented in the R-packages \texttt{mdmb} \citep{mdmb} for
imputation of missing data and \texttt{synthpop} \citep{synthpop} for
imputation of synthetic data.

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(mice)}
\NormalTok{imp }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(data.to.synthesize, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{1}\NormalTok{,}
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{,}
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(data.to.synthesize),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(data.to.synthetize)))}
\end{Highlighting}
\end{Shaded}

@@

However, one of the methods that has shown a reasonable performance over
multiple simulations uses FCS in combination with classification and
regression trees
\citep[CART;][]{breiman_cart_1984, reiter_cart_2005, burgette_reiter_cart_2010, doove_buuren_recursive_2014, raab_practical_2016}.
Due to its non-parametric nature, CART allows to model more complex
patterns (e.g., interactions and non-linear relationships) in the data.
Loosely speaking, CART sequentially splits the predictor space into
non-overlapping regions in such a way that the within-region variance is
as small as possible after every split. As such, CART does not impose
any parametric distribution on the data, making it a widely applicable
method that allows for a large variety of relationships within the data
\citep{islr_2013}. Given these appealing characteristics and the call
for the use of flexible methods when multiply imputing data, we will
focus our illustrations and evaluations of \texttt{mice} to method
\texttt{mice.impute.cart()}, realized by:

@@

In a nutshell, the above code shows the simplicity of creating
\(m = 10\) synthetic data sets using \texttt{mice}. In practice,
however, one should take some additional complicating factors into
account. For example, one should account for deterministic relations in
the data. Additionally, relations between variables may be described
best using a different model than \texttt{CART}. Such factors are data
dependent and should be considered by the imputer. In the next section,
we will describe how the \texttt{mice::boys} data
{[}fredriks\_boys\_2000{]} can be adequately synthesized. Additionally,
we will show through simulations that this approach yields valid
inferences.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{We are grateful to Mirthe Hendriks and Stijn van den
Broek for replicating the simulation study on an independent data set.
Additionally, we thank Hanne Oberman and two anonymous reviewers for
their stimulating comments.}

>>>>

Anonymizing research data might seem a quick and appealing approach to
limit the unique identification of participants. However, this approach
is not sufficient to fulfill contemporary privacy and confidentiality
requirements \citep{ohm_broken_2009, national_putting_2007}. Over the
years, several other techniques have been used to increase the
confidentiality of research data, such as categorizing continuous
variables, top coding values above an upper bound or adding random noise
to the observed values \citep{drechsler_synthetic_2011}. However, these
methods may distort the true data relation between variables, thereby
reducing the data quality and the scientific returns for re-using the
same data for further research.

@@

Generally, there are three ways to impute data: joint modeling,
sequential modeling and fully conditional specification
\citep{murray_multiple_2018, ludtke2020regression, grund2021using}. With
joint modeling, the entire joint distribution of the data is specified
as a single multivariate distribution, and the imputations are drawn
from this model. This strategy is implemented in, among others, the
\texttt{R}-package \texttt{jomo} \citep{jomo} to deal with missing data.
However, when the structure of the data increases in complexity,
specifying a single multivariate distribution that would fit the
observed data can become challenging. A more flexible solution has been
proposed in sequential regression imputation, in which the multivariate
distribution is separated into univariate conditional distributions.
Every single variable is then imputed conditionally on a model
containing only the variables in the sequence located before the
variable to be imputed. This approach has been implemented in the
R-packages \texttt{mdmb} \citep{mdmb} for imputation of missing data and
\texttt{synthpop} \citep{synthpop} for imputation of synthetic data.

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(mice)}
\NormalTok{syn }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(data.to.synthesize, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{1}\NormalTok{,}
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{,}
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(data.to.synthesize),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(data.to.synthetize)))}
\end{Highlighting}
\end{Shaded}

@@

However, one of the methods that has shown a reasonable and consistent
performance over multiple simulations using FCS in combination with
classification and regression trees
\citep[CART;][]{breiman_cart_1984, reiter_cart_2005, burgette_reiter_cart_2010, doove_buuren_recursive_2014, raab_practical_2016}.
Due to its non-parametric nature, CART allows to model more complex
patterns (e.g., interactions and non-linear relationships) in the data.
Loosely speaking, CART sequentially splits the predictor space into
non-overlapping regions in such a way that the within-region variance is
as small as possible after every split. As such, CART does not impose
any parametric distribution on the data, making it a widely applicable
method that allows for a large variety of relationships within the data
\citep{islr_2013}. Given these appealing characteristics and the call
for the use of flexible methods when multiply imputing data, we will
focus our illustrations and evaluations of \texttt{mice} to method
\texttt{mice.impute.cart()}, realized by:

@@

In a nutshell, the above code shows the simplicity of creating
\(m = 10\) synthetic data sets using \texttt{mice}. In practice,
however, one should take some additional complicating factors into
account. For example, one should account for deterministic relations in
the data. Additionally, relations between variables may be described
best using a different model than \texttt{CART}. Such factors are data
dependent and should be considered by the imputer. In the next section,
we will describe how the \texttt{mice::boys} data
\citep{fredriks_boys_2000} can be adequately synthesized. Additionally,
we will show through simulations that this approach yields valid
inferences.

@@

\begin{verbatim}
## Loading required package: lattice
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'caret'
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:future':
## 
##     cluster
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:purrr':
## 
##     lift
\end{verbatim}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{We are grateful to Mirthe Hendriks and Stijn van den
Broek for replicating the simulation study on an independent data set.
Additionally, we thank Hanne Oberman and two anonymous reviewers for
their helpful and insightful comments.}
